Measuring Perceptions and Preferences for Meritocracy

### 2023-02-02 paper-blinded.Rmd
ajustes layout

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
# author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies. The belief in meritocracy has been suggested as one of the factors that explain the persistence of inequality, as it would legitimize economic disparities based on differences in effort and talent. Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualizations and measures that make the findings and their interpretation rather inconsistent. Most of the studies in the area have relied upon secondary data from different surveys to operationalize meritocracy, with a wide variation in the use and interpretation of the same measuring items. Based on a critical analysis of concepts and operationalizations of meritocracy in survey research, we propose a conceptual and measurement framework for perceptions and preferences for meritocracy and non-meritocracy in a minimalistic version (few items as possible), in order to be considered in comparative survey research studies. The proposal is first tested through confirmatory factor analysis (CFA) using available ISSP (International Social Survey Programme) inequality module items. Based on these results, in a second study we propose an  improved scale that better suits the perceptions/preferences distinction and that is tested  in a Chilean sample (N=2,141). The results of the CFA estimation give evidence of a better fit for the new scale. Our discussion highlights the importance of considering different dimensions in the study of meritocracy, both conceptually and empirically."
output:
  bookdown::pdf_document2:
     template: null
     toc: false
     number_sections: false
  bookdown::word_document2:
    number_sections: false
  bookdown::html_document2:
    number_sections: false
linkcolor: blue
bibliography:
  - input/bib/meritocracy.bib
csl: input/bib/apa7-singlespace.csl
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage[figurename=Fig.]{caption}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
always_allow_html: true # for word output
fontsize: 10pt
---

@@

```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo=FALSE, warning = FALSE,message = FALSE, cache = FALSE,out.width = '85%',fig.pos= "H", fig.align = 'center')
 # knitr::opts_knit$set(base.url = "../") #relative path for .html output file
 # knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
 options(scipen=999)
 options(kableExtra.auto_format = FALSE)
 rm(list=ls())
 options(knitr.kable.NA = '')
 options(knitr.graphics.error = FALSE)
 Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
```

@@

```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Second-order confirmatory factor analysis model using ISSP indicators of Perceptions and Preferences for Meritocracy"}
 knitr::include_graphics('output/images/Fig4.cfa_issp_study1.PNG')
```

@@

```{r appmod, echo=FALSE, fig.cap = "Survey flow", fig.align='center'}
 knitr::include_graphics('output/images/Fig5.app_mod.png')
```

@@

A pre-registration was made in the OSF platform, available at the following link: [https://osf.io/z45y2](https://osf.io/z45y2). This pre-registration includes the hypotheses regarding the four-dimensional conceptual model underlying the scale, the variable measurement levels, the statistical tests to be performed with their respective evaluation parameters, and other important aspects of the research design.

@@

```{r corpoly, echo=FALSE, fig.cap = "Perceptions and preferences for meritcoracy items' polychoric correlations", fig.align='center'}
  # dev.copy(png,'output/images/Fig7.plotcor-study2.png',width = 800, height = 600) # save image
  if (!require("pacman")) install.packages("pacman")
  load(file = "input/data/proc/data_s2.RData")
  pacman::p_load(dplyr, corrplot, lavaan)
  # generate polychoric matrix
  cor_s2<- data_s2 %>%
    dplyr::select(perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw) %>%
    lavaan::lavCor(., ordered=names(.))

@@

```{r meas02, echo=FALSE, fig.cap = "Confirmatory factor analysis of the Perceptions and Preferences for Meritocracy Scale", fig.align='center'}
 # estimation for diagram: summary(fit_cfa3, standardized = TRUE, fit.measures=TRUE)
 knitr::include_graphics('output/images/Fig8.cfa_randomized_study2.png')
```

>>>>

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '2.0'
link-citations: yes
# author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
# abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies. The belief in meritocracy has been suggested as one of the factors that explain the persistence of inequality, as it would legitimize economic disparities based on differences in effort and talent. Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualizations and measures that make the findings and their interpretation rather inconsistent. Most of the studies in the area have relied upon secondary data from different surveys to operationalize meritocracy, with a wide variation in the use and interpretation of the same measuring items. Based on a critical analysis of concepts and operationalizations of meritocracy in survey research, we propose a conceptual and measurement framework for perceptions and preferences for meritocracy and non-meritocracy in a minimalistic version (few items as possible), in order to be considered in comparative survey research studies. The proposal is first tested through confirmatory factor analysis (CFA) using available ISSP (International Social Survey Programme) inequality module items. Based on these results, in a second study we propose an  improved scale that better suits the perceptions/preferences distinction and that is tested  in a Chilean sample (N=2,141). The results of the CFA estimation give evidence of a better fit for the new scale. Our discussion highlights the importance of considering different dimensions in the study of meritocracy, both conceptually and empirically."
output:
  bookdown::pdf_document2:
     template: null
     toc: false
     number_sections: false
     keep_text: true
  # bookdown::word_document2:
  #   number_sections: false
  # bookdown::html_document2:
  #   number_sections: false
linkcolor: blue
bibliography:
  - input/bib/meritocracy.bib
csl: input/bib/asa6.csl
editor_options:
  chunk_output_type: console
geometry: margin=1in
header-includes:
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
#  - \usepackage[figurename=Fig.]{caption}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
always_allow_html: true # for word output
fontsize: 12pt
papersize: a4
---

@@

```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo=FALSE, warning = FALSE,message = FALSE, cache = TRUE,out.width = '85%',fig.pos= "H", fig.align = 'center')
 # knitr::opts_knit$set(base.url = "../") #relative path for .html output file
 # knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
 options(scipen=999)
 options(kableExtra.auto_format = FALSE)
 rm(list=ls())
 options(knitr.kable.NA = '')
 options(knitr.graphics.error = FALSE)
 Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
```

@@

```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Second-order confirmatory factor analysis model using ISSP indicators of Perceptions and Preferences for Meritocracy"}
 knitr::include_graphics('output/images/figure4.PNG')
```

@@

```{r appmod, echo=FALSE, fig.cap = "Survey flow", fig.align='center'}
 knitr::include_graphics('output/images/figure6.png')
```

@@

A pre-registration was made in the OSF platform, available at the following link: [https://osf.io/z45y2?](https://osf.io/z45y2/?view_only=adcd496887f2471597b1d6ffb46e492d). This pre-registration includes the hypotheses regarding the four-dimensional conceptual model underlying the scale, the variable measurement levels, the statistical tests to be performed with their respective evaluation parameters, and other important aspects of the research design.

@@

```{r corpoly, echo=FALSE, fig.cap = "Perceptions and preferences for meritocracy items' polychoric correlations", fig.align='center'}
  # dev.copy(png,'output/images/Fig7.plotcor-study2.png',width = 800, height = 600) # save image
  if (!require("pacman")) install.packages("pacman")
  load(file = "input/data/proc/data_s2.RData")
  pacman::p_load(dplyr, corrplot, lavaan)
  # generate polychoric matrix
  cor_s2<- data_s2 %>%
    dplyr::select(perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw) %>%
    lavaan::lavCor(., ordered=names(.))

@@

```{r meas02, echo=FALSE, fig.cap = "Confirmatory factor analysis of the Perceptions and Preferences for Meritocracy Scale", fig.align='center'}
 # estimation for diagram: summary(fit_cfa3, standardized = TRUE, fit.measures=TRUE)
 knitr::include_graphics('output/images/figure9.png')
```
